---
title: "Review Data"
author: "Sophia Deng"
date: "2024-04-28"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5, tidy = TRUE)

library(geomtextpath) # for geom_textvline
library(lubridate)
library(mgcv)
library(tidyverse)

source("ggplot_settings.R")
theme_set(theme_custom())
```

## Split up combined reviews into 3 categories

```{r eval=F}
load("../review_data/combined_reviews.RData")

asian_reviews <- combined_reviews %>% filter(str_detect(type, "asian"))
asian_reviews <- na.omit(asian_reviews)
write.csv(asian_reviews, "../review_data/all_asian_reviews.csv")

pizza_reviews <- combined_reviews %>% filter(str_detect(type, "pizza"))
pizza_reviews <- na.omit(pizza_reviews)
write.csv(pizza_reviews, "../review_data/all_pizza_reviews.csv")

mexican_reviews <- combined_reviews %>% filter(str_detect(type, "mexican"))
mexican_reviews <- na.omit(mexican_reviews)
write.csv(mexican_reviews, "../review_data/all_mexican_reviews.csv")
```

## Plot average sentiment of Asian reviews

```{r}
asian_sentiment_reviews <- read_csv("../sentiment_data/asian_sentiment.csv") |>
  mutate(date = as.Date(paste(year, month, 01), "%Y %m %d")) |>
  filter(time <= as.Date(paste(2021, 06, 01), "%Y %m %d"))

# Calculate average sentiment per month
asian_avg_sentiment <- aggregate(polarity ~ date, data = asian_sentiment_reviews, FUN = mean)

# Plot the graph
ggplot(asian_avg_sentiment, aes(x = date, y = as.numeric(polarity))) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "loess", se = FALSE) +
  labs(x = "Date", y = "Average Sentiment", title = "Average Sentiment per Month for Asian restaurants")
```

## Plot average sentiment of pizza reviews

```{r}
pizza_sentiment_reviews <- read_csv("../sentiment_data/pizza_sentiment.csv") |>
  mutate(date = as.Date(paste(year, month, 01), "%Y %m %d")) |>
  filter(time <= as.Date(paste(2021, 06, 01), "%Y %m %d"))

# Calculate average sentiment per month
pizza_avg_sentiment <- aggregate(polarity ~ date, data = pizza_sentiment_reviews, FUN = mean)

# Plot the graph
ggplot(pizza_avg_sentiment, aes(x = date, y = as.numeric(polarity))) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "loess", se = FALSE) +
  labs(x = "Date", y = "Average Sentiment", title = "Average Sentiment per Month for pizza restaurants")
```

## Plot average sentiment of Mexican reviews

```{r}
mexican_sentiment_reviews <- read_csv("../sentiment_data/mexican_sentiment.csv") |>
  mutate(date = as.Date(paste(year, month, 01), "%Y %m %d")) |>
  filter(time <= as.Date(paste(2021, 06, 01), "%Y %m %d"))

# Calculate average sentiment per month
mexican_avg_sentiment <- aggregate(polarity ~ date, data = mexican_sentiment_reviews, FUN = mean)

# Plot the graph
ggplot(mexican_avg_sentiment, aes(x = date, y = as.numeric(polarity))) +
  geom_point() +
  geom_smooth(aes(group = 1), method = "loess", se = FALSE) +
  labs(x = "Date", y = "Average Sentiment", title = "Average Sentiment per Month")
```

```{r}
# REMOVING LAST 3 MONTHS OF 2021 DATA

# create combined sentiment reviews
combined_reviews <-
  bind_rows(asian_sentiment_reviews, mexican_sentiment_reviews, pizza_sentiment_reviews) |>
  mutate(date = as.Date(paste(year, month, 01), "%Y %m %d")) |>
  filter(time <= as.Date(paste(2021, 06, 01), "%Y %m %d"))
```

## Plot average polarity 

```{r}
covid_start <- as.Date(paste(2020, 03, 15), "%Y %m %d")
stopah_start <- as.Date(paste(2021, 03, 11), "%Y %m %d")

combined_reviews |>
  mutate(type = factor(str_to_title(type), levels = c("Asian", "Mexican", "Pizza"))) |>
  group_by(type, month, year) |>
  summarise(mean_polarity = mean(polarity)) |>
  mutate(date = as.Date(paste(year, month, 01), "%Y %m %d")) |>
  filter(date <= as.Date(paste(2021, 06, 01), "%Y %m %d")) |>
  ggplot(aes(x = date, y = mean_polarity, color = type)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("lightpink", "steelblue", "darkseagreen")) +
  geom_vline(xintercept = covid_start, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = stopah_start, color = "dodgerblue2", lty = "dashed") +
  labs(title = "Average polarity over time for Asian, Mexican and pizza restaurants", color = "Cuisine", x = "Date", y = "Average polarity")

# ggsave(filename = "../figures/average_polarity.png", width = 8, height = 5)
```

## Plot average polarity with smoothed line

```{r}
avg_sentiment <- bind_rows(pizza_avg_sentiment, asian_avg_sentiment, mexican_avg_sentiment) |>
  mutate(type = rep(c("pizza", "asian", "mexican"), each = 29))

ggplot(avg_sentiment) +
  geom_point(aes(x = date, y = polarity, color = type)) +
  geom_smooth(aes(x = date, y = polarity, color = type), method = "loess", se = F) +
  scale_color_manual(values = c("lightpink", "steelblue", "darkseagreen"), labels = c("Asian", "Mexican", "Pizza")) +
  geom_vline(xintercept = covid_start, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = stopah_start, color = "dodgerblue2", lty = "dashed") +
  labs(x = "Date", y = "Average polarity", title = "Average polarity over time for Asian, Mexican and pizza restaurants", color = "Cuisine")
# ggsave("../figures/average_monthly_sentiment.png", width = 8, height = 5)
```

# GAM with Date 0 as 3/11/2020 (start of covid)

```{r}
combined_reviews <- combined_reviews |> mutate(covid_date = as.numeric(lubridate::as_date(time) - lubridate::as_date("2020-03-11")))
asian_sentiment <- combined_reviews %>% filter(type == "asian")
mexican_sentiment <- combined_reviews %>% filter(type == "mexican")
pizza_sentiment <- combined_reviews %>% filter(type == "pizza")

# using gam model with te (tensor) which allows for an interaction
gam_asian <- mgcv::gam(polarity ~ 1 + te(rating, covid_date, k = c(4, 10)), data = asian_sentiment)
summary(gam_asian)

gam_mexican <- mgcv::gam(polarity ~ 1 + te(rating, covid_date, k = c(4, 10)), data = mexican_sentiment)
summary(gam_mexican)

gam_pizza <- mgcv::gam(polarity ~ 1 + te(rating, covid_date, k = c(4, 10)), data = pizza_sentiment)
summary(gam_pizza)
```

# Comparing 1 star reviews

```{r}
one_star_df <- data.frame(covid_date = -435:446, rating = 1)

asian_polarity <- predict(gam_asian, newdata = one_star_df)
mexican_polarity <- predict(gam_mexican, newdata = one_star_df)
pizza_polarity <- predict(gam_pizza, newdata = one_star_df)

one_star_preds <- bind_rows(
  bind_cols(one_star_df, polarity = asian_polarity, type = "asian"),
  bind_cols(one_star_df, polarity = mexican_polarity, type = "mexican"),
  bind_cols(one_star_df, polarity = pizza_polarity, type = "pizza")
)

ggplot(data = one_star_preds, aes(x = covid_date, y = polarity, color = type)) +
  geom_line(lwd = 0.1) +
  geom_point() +
  scale_color_manual(values = c("lightpink", "steelblue", "darkseagreen")) +
  geom_vline(xintercept = 0, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = 370, color = "dodgerblue2", lty = "dashed") +
  labs(title = "Polarity over time for 1 star reviews", color = "Cuisine", x = "Days from US COVID-19 Lockdown", y = "Polarity") +
  theme(plot.title = element_text(hjust = 0.5))
# ggsave(filename = "../figures/polarity_for_1_star.png", width = 8, height = 5)
```

# Comparing 5 star reviews

```{r}
all_stars_df <- data.frame(covid_date = rep(-435:446, each = 5), rating = 1:5)

asian_polarity <- predict(gam_asian, newdata = all_stars_df)
mexican_polarity <- predict(gam_mexican, newdata = all_stars_df)
pizza_polarity <- predict(gam_pizza, newdata = all_stars_df)

all_stars_preds <- bind_rows(
  bind_cols(all_stars_df, polarity = asian_polarity, type = "asian"),
  bind_cols(all_stars_df, polarity = mexican_polarity, type = "mexican"),
  bind_cols(all_stars_df, polarity = pizza_polarity, type = "pizza")
)

all_stars_preds |>
  filter(type == "pizza") |>
  ggplot(aes(x = covid_date, y = polarity, color = as.character(rating))) +
  geom_line(lwd = 1) +
  geom_vline(xintercept = 0, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = 370, color = "dodgerblue2", lty = "dashed") +
  labs(title = "Polarity over time for pizza reviews", color = "Rating", x = "Days from US COVID-19 Lockdown", y = "Polarity")

all_stars_preds |>
  filter(type == "asian") |>
  ggplot(aes(x = covid_date, y = polarity, color = as.character(rating))) +
  geom_line(lwd = 1) +
  geom_vline(xintercept = 0, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = 370, color = "dodgerblue2", lty = "dashed") +
  labs(title = "Polarity over time for Asian reviews", color = "Rating", x = "Days from US COVID-19 Lockdown", y = "Polarity")
# ggsave("../figures/polarity_asian_1_to_5_stars.png", width = 8, height = 5)

all_stars_preds |>
  filter(type == "mexican") |>
  ggplot(aes(x = covid_date, y = polarity, color = as.character(rating))) +
  geom_line(lwd = 1) +
  geom_vline(xintercept = 0, color = "indianred2", lty = "dashed") +
  geom_vline(xintercept = 370, color = "dodgerblue2", lty = "dashed") +
  labs(title = "Polarity over time for Mexican reviews", color = "Rating", x = "Days from US COVID-19 Lockdown", y = "Polarity")
```


### Violin plots

```{r}
# polarity vs rating
ggplot(data = asian_sentiment, aes(x = as.character(rating), y = polarity)) +
  geom_violin()

ggplot(data = mexican_sentiment, aes(x = as.character(rating), y = polarity)) +
  geom_violin()

ggplot(data = pizza_sentiment, aes(x = as.character(rating), y = polarity)) +
  geom_violin()
```


## Actual polarity vs date, coloured by rating

```{r}
# polarity vs date, color by rating

ggplot(data = asian_sentiment, aes(x = covid_date, y = polarity)) +
  geom_smooth(aes(color = as.factor(rating)), method = "loess", se = FALSE) +
  labs(title = "Actual polarity across time for Asian restaurants")

ggplot(data = mexican_sentiment, aes(x = covid_date, y = polarity)) +
  geom_smooth(aes(color = as.factor(rating)), method = "loess", se = FALSE) +
  labs(title = "Actual polarity across time for Mexican restaurants")

ggplot(data = pizza_sentiment, aes(x = covid_date, y = polarity)) +
  geom_smooth(aes(color = as.factor(rating)), method = "loess", se = FALSE) +
  labs(title = "Actual polarity across time for pizza restaurants")
```


Projected range along one variable (date), but fixed ratings (between 1 to 5) for all observations. We plotted the response of the model for fixed values of the date. 

For a random forest: we have a median value. Can do this for up to two variables: interaction patterns, plotting by heatmap: color boxes by model output (vary two variables on x and y axis, are there hotspots for large or small model outputs)
