---
title: "loading review data"
format: html
---

```{r}
library(jsonlite)
library(tidyverse)
library(foreach)

setwd("~/Documents/GitHub/small-business-sentiment-analysis/scripts/")
```


# Load data

The files are from https://datarepo.eng.ucsd.edu/mcauley_group/gdrive/googlelocal/

in the format where

user_id - ID of the reviewer
name - name of the reviwer
time - time of the review (unix time)
rating - rating of the business
text - text of the review
pics - pictures of the review
resp - business response to the review including unix time and text of the response
gmap_id - ID of the business

and crosslink with gmap_id in metadata where

name - name of the business
address - address of the business
gmap_id - ID of the business
description - description of the business
latitude - latitude of the business
longitude - longitude of the business
category - category of the business
avg_rating - average rating of the business
num_of_reviews - number of reviews
price - price of the business
hours - open hours
MISC - MISC information
state - the current status of the business (e.g., permanently closed)
relative_results - relative businesses recommended by Google
url - URL of the business


```{r eval=F}
meta_file <- "../review_data/meta-New_York.json"
meta_json <- purrr::map(readLines(meta_file), jsonlite::fromJSON, .progress = "loading metadata")

l <- map(meta_json, function(x) {
  cbind(
    ifelse(is.null(x$name), NA, x$name),
    x$gmap_id,
    ifelse(is.null(x$address), NA, x$address),
    ifelse(is.null(x$category), NA, paste(x$category, collapse = ", "))
  )
})
metadata <- data.frame(matrix(unlist(l), ncol = 4, byrow = TRUE))
colnames(metadata) <- c("name", "gmap_id", "address", "category")
metadata <- tibble(metadata) |> distinct()
```

```{r eval=F}
filepaths <- list.files(path = "../review_data/review_data_split")

foreach(i = seq(1, 100, by = 5)) %do% {
  reviews_by_5 <- foreach(current_file = filepaths[i:(i + 4)], .combine = "rbind") %do% {
    review_json <- purrr::map(readLines(paste0("../review_data/review_data_split/", current_file)), jsonlite::fromJSON, .progress = paste("loading file", current_file))
    l <- map(review_json, function(x) {
      ƒ
      cbind(x$time, ifelse(is.null(x$rating), NA, x$rating), ifelse(is.null(x$text), NA, x$text), x$gmap_id)
    })
    r <- data.frame(matrix(unlist(l), ncol = 4, byrow = TRUE))
    colnames(r) <- c("time", "rating", "text", "gmap_id")
    r
  }
  reviews_by_5 <- reviews_by_5 |> distinct()
  save(reviews_by_5, file = paste0("reviews_", ceiling(i / 5), ".RData"))
  rm(reviews_by_5)
}
```

